---
title: "lifestage-qPCR"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/project-gigas-carryover")
getwd()
```

### loading libraries

```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')

library("tidyverse")
library("ggplot2")
```

### set variables (unsure of what this is for atm)

```{r}
seed.control <- c("223", "243", "244")
seed.treated <- c("200", "257", "285")
spat.control <- c("206", "282", "284", "289")
spat.treated <- c("220", "226", "242", "253", "296", "298")

# Combine vectors into a list
vector_list <- list(seed.control = seed.control,
                    seed.treated = seed.treated,
                    spat.control = spat.control,
                    spat.treated = spat.treated)
```

### functions

```{r}
calculate_delta_Cq <- function(df) {
  df <- df %>%
    group_by(Sample) %>%
    mutate(delta_Cq = Cq.Mean - Cq.Mean[Target == "GAPDH"]) %>%
    ungroup()
  
  return(df)
}
```

### Read in files

```{r}
# Set the directory where your CSV files are located
cqs_directory <- "project-gigas-carryover/lifestage_carryover/data/qPCR/Cq"

# Get a list of all CSV files in the directory with the naming structure "*Cq_Results.csv"
cq_file_list <- list() # Initialize list
cq_file_list <- list.files(path = cqs_directory, pattern = "Cq_Results\\.csv$", full.names = TRUE)

# Initialize an empty list to store the data frames
data_frames <- list()

# Loop through each file and read it into a data frame, then add it to the list
for (file in cq_file_list) {
  data <- read.csv(file, header = TRUE)
  data_frames[[file]] <- data
}

# Combine all data frames into a single data frame
combined_df <- bind_rows(data_frames, .id = "data_frame_id")

# Convert Sample column to character type
combined_df <- combined_df %>%
  mutate(Sample = as.character(Sample))

str(combined_df)
```

### unique samples by target

```{r}
# Group by Sample and Target, then summarize to get unique rows for each sample
aggregated_df <- combined_df %>%
  group_by(Sample, Target) %>%
  summarize(Cq.Mean = mean(Cq.Mean, na.rm = TRUE)) %>%
  ungroup()

str(aggregated_df)

```

### Add life stage and treatment cols

```{r}
# Initialize new columns
aggregated_df <- aggregated_df %>%
  mutate(life.stage = NA_character_,
         treatment = NA_character_)

# Loop through each vector
for (vec_name in names(vector_list)) {
  vec <- vector_list[[vec_name]]
  stage <- strsplit(vec_name, "\\.")[[1]][1]
  treatment <- strsplit(vec_name, "\\.")[[1]][2]
  
  # Loop through each row in aggregated_df
  for (i in 1:nrow(aggregated_df)) {
    sample <- aggregated_df$Sample[i]
    
    # Check if sample is in the vector
    if (sample %in% vec) {
      # Update life.stage and treatment columns
      aggregated_df$life.stage[i] <- stage
      aggregated_df$treatment[i] <- treatment
    }
  }
}

str(aggregated_df)
```

### Delta Cq to Normalizing Gene

```{r}
# Calculate delta Cq by subtracting GAPDH Cq.Mean from each corresponding Sample Cq.Mean
delta_Cq_df <- calculate_delta_Cq(aggregated_df)

str(delta_Cq_df)
```

### T-tests
#### This is taken directly from [sam's notebook](https://robertslab.github.io/sams-notebook/posts/2024/2024-03-27-qPCR-Analysis---C.gigas-Lifestages-Carryover-from-20240325/#unique-samples-by-target) . His code surround the "asterisk" is clever and useful.

```{r}


# Filter out groups with missing life.stage or Target
# Caused by NTCs
# Also removes normalizing gene(s)
delta_Cq_df_filtered <- delta_Cq_df %>%
  filter(!is.na(life.stage), !is.na(Target), Target != "GAPDH")

# Perform t-test for each Target within life.stage
t_test_results <- delta_Cq_df_filtered %>%
  group_by(life.stage, Target) %>%
  summarise(
    t_test_result = list(t.test(delta_Cq ~ treatment))
  ) %>%
  ungroup()

# Extract t-test statistics
t_test_results <- t_test_results %>%
  mutate(
    estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
    p_value = sapply(t_test_result, function(x) x$p.value)
  ) %>% 
  select(!t_test_result)

# Add asterisk information to data frame
# Useful for plotting
t_test_results$asterisk <- ifelse(t_test_results$p_value < 0.05, "*", "")

head(t_test_results)
```

### Delta Cq Boxplots !!
#### seed

```{r}
# Filter delta_Cq_df_filtered for seed life stage
seed_delta_Cq_df <- delta_Cq_df_filtered %>%
  filter(life.stage == "seed")

# Create the box plot
boxplot <- ggplot(seed_delta_Cq_df, aes(x = Target, y = delta_Cq, colour = treatment)) +
  geom_boxplot(aes(colour = treatment))+
  theme(legend.position = "right")+
  labs(x = "Target", y = "Delta Cq", title = "Seed Gene Expression", colour = "Treatment")+
  scale_color_manual(values = c("treated" = "red3", "control" = "steelblue"))+
  theme_minimal()

# Add asterisks
boxplot  +
   annotate("text", x = t_test_results$Target, y = Inf, label = t_test_results$asterisk,
           vjust = -0.5, size = 4, color = "orange")
```

#### spat
```{r}
# Filter delta_Cq_df_filtered for seed life stage
spat_delta_Cq <- delta_Cq_df_filtered %>%
  filter(life.stage == "spat")

# Calculate the maximum delta_Cq for each Target
max_delta_Cq_by_target <- spat_delta_Cq %>%
  group_by(Target) %>%
  summarise(max_delta_Cq = max(delta_Cq, na.rm = TRUE))

# Merge t_test_results with max_delta_Cq_by_target to get the maximum delta_Cq for each Target
t_test_results_with_max_delta_Cq <- merge(t_test_results, max_delta_Cq_by_target, by = "Target")

# Create the box plot
boxplot <- ggplot(spat_delta_Cq, aes(x = Target, y = delta_Cq, colour = treatment)) +
  geom_boxplot(aes(colour = treatment))+
  theme(legend.position = "right")+
  labs(x = "Target", y = "Delta Cq", title = "Spat Gene Expression", colour = "Treatment")+
  scale_color_manual(values = c("treated" = "red3", "control" = "steelblue"))+
  theme_minimal()

# Add asterisks for statistically significant differences (t-test)
boxplot +
 annotate("text", x = t_test_results_with_max_delta_Cq$Target, 
           y = t_test_results_with_max_delta_Cq$max_delta_Cq + 0.5, 
           label = t_test_results_with_max_delta_Cq$asterisk,
           vjust = -0.5, size = 10, color = "orange")
```

## Delta delta Cq

### Add treatment and life stage

```{r add-treatment-lifestage}
# Initialize empty vectors to store life.stage and treatment
life_stage <- character(nrow(combined_df))
treatment <- character(nrow(combined_df))

# Loop through each row of combined_df
for (i in 1:nrow(combined_df)) {
  sample_id <- combined_df$Sample[i]
  
  # Check if the sample_id is present in any of the vectors
  found <- FALSE
  for (vec_name in names(vector_list)) {
    if (sample_id %in% vector_list[[vec_name]]) {
      # If present, extract life.stage and treatment from the vector name
      parts <- strsplit(vec_name, "\\.")[[1]]
      life_stage[i] <- parts[1]
      treatment[i] <- parts[2]
      found <- TRUE
      break  # Exit loop once found
    }
  }
  
  # If sample_id is not found in any vector, assign NA to both life.stage and treatment
  if (!found) {
    life_stage[i] <- NA
    treatment[i] <- NA
  }
}

# Add life.stage and treatment columns to combined_df
combined_df <- combined_df %>%
  mutate(life.stage = life_stage,
         treatment = treatment)

# Filter out rows where life.stage is NA
combined_df_filtered <- combined_df %>% 
  filter(!is.na(life.stage))

str(combined_df_filtered)
```


### Mean Cqs per gene per treatment per life stage
```{r mean-Cqs}
# Group by life.stage, treatment, and Target, then calculate the mean Cq
mean_Cq_df <- combined_df_filtered %>%
  group_by(life.stage, treatment, Target) %>%
  summarise(mean_Cq = mean(Cq, na.rm = TRUE))
```

### Delta Cqs

```{r delta-Cqs-gene-level}
# Calculate delta Cq
combined_df_with_delta_Cq <- mean_Cq_df %>%
  group_by(life.stage, treatment) %>%
  mutate(delta_Cq = mean_Cq - mean(mean_Cq[Target == "GAPDH"])) %>%
  ungroup() %>% 
  filter(Target != "GAPDH")
```

### Delta delta Cq

```{r delta-delta-Cq}
# Calculate delta_delta_Cq
delta_delta_Cq_df <- combined_df_with_delta_Cq %>%
  group_by(life.stage, Target) %>%
  summarize(delta_delta_Cq = delta_Cq[treatment == "treated"] - delta_Cq[treatment == "control"])

```

### Calculate the fold change for each Target
```{r fold-change}
delta_delta_fold_change <- delta_delta_Cq_df %>%
  mutate(fold_change = 2^(-delta_delta_Cq)) %>% 
  distinct(Target, fold_change)

str(delta_delta_fold_change)
```

### Plot - Seed Fold Change

```{r plot-seed-fold-change}
#| label: fig-bar-plots-seed-fold-change
#| fig-cap: "Bar plots showing fold change in expression (2^(-delta delta Cq)) in seed."
#| fig-alt: "Bar plots showing fold change in expression (2^(-delta delta Cq)) in seed. HSP90 is the only gene which is more highly expressed in treatment relative to controls in seed."


# Filter delta_delta_fold_change for seed life stage
seed_df <- delta_delta_fold_change %>%
  filter(life.stage == "seed")

# Create bar plot for seed life stage
# Create the plot with fold changes relative to baseline of 1
seed_plot <- ggplot(seed_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Seed Fold Change") +
  scale_y_continuous(limits = c(-1, 8))

# Display plot
seed_plot

```

### Plot - Spat Fold Change
```{r plot-spat-fold-change}
#| label: fig-bar-plots-spat-fold-change
#| fig-cap: "Bar plots showing fold change in expression (2^(-delta delta Cq)) in spat."
#| fig-alt: "Bar plots showing fold change in expression (2^(-delta delta Cq)) in spat. HSP90 is the only gene which is more highly expressed in treatment relative to controls in spat."
#| 
# Filter delta_delta_fold_change for spat life stage
spat_df <- delta_delta_fold_change %>%
  filter(life.stage == "spat")

# Create bar plot for spat life stage
# Create the plot with fold changes relative to baseline of 1
spat_plot <- ggplot(spat_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Spat Fold Change") +
  scale_y_continuous(limits = c(-1, 8))

# Display plot
spat_plot
```
