---
title: "01.01-qPCR"
author: "Sam White"
date: "2024-03-27"
output: 
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  github_document:
    toc: true
    number_sections: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
bibliography: references.bib
link-citations: true
---

# Load libraries
```{r load-libraries}
library("tidyverse")
library("ggplot2")
```

# Set variables

## Set sample groups
```{r set-sample-groups}
seed.control=c("223", "243", "244")
seed.treated=c("200", "257", "285")
spat.control=c("206", "282", "284", "289")
spat.treated=c("220", "226", "242", "253", "296", "298")
juvenile.control.ambient=c("18", "57", "65", "75", "79", "104", "110", "123", "125", "171", "175", "205", "238", "273", "279", "293", "317")
juvenile.control.high=c("12", "39", "43", "49", "71", "130", "141", "146", "150", "170", "195", "297", "301", "324", "351", "355", "371")
juvenile.treated.ambient=c("1", "34", "64", "83", "98", "147", "152", "158", "162", "169", "188", "271", "295", "310", "357", "361", "381")
juvenile.treated.high=c("28", "53", "61", "73", "81", "106", "109", "139", "149", "173", "181", "213", "290", "302", "311", "364", "392")
adult.control.ambient=c("3", "5", "13*", "16", "17", "80", "87", "94", "148", "159", "179", "180", "250", "258", "268", "312", "326", "330", "334", "346", "360", "377", "379", "386")
adult.control.high=c("20", "23", "26", "32", "33", "67", "70", "90", "107", "132", "135", "157", "166", "186", "207", "215", "248", "316", "341", "344", "349", "382", "394", "395")
adult.treated.ambient=c("7", "31", "35", "37", "41", "54", "84", "100", "112", "116", "118", "133", "154", "199", "203", "204", "208", "219", "294", "318", "339", "353", "363", "378")
adult.treated.high=c("21", "22", "45", "82", "85", "91", "95", "99", "103", "108", "117", "127", "165", "185", "190", "196", "232", "237", "245", "263", "276", "306", "343", "374")

# Combine vectors into lists
# Used for adding treatment info and/or subsetting downstream
seed.spat_vector_list <- list(seed.control = seed.control,
                    seed.treated = seed.treated,
                    spat.control = spat.control,
                    spat.treated = spat.treated)

adult.juvenile_vector_list <- list(juvenile.control.ambient = juvenile.control.ambient,
                                   juvenile.control.high = juvenile.control.high,
                                   juvenile.treated.ambient = juvenile.treated.ambient,
                                   juvenile.treated.high = juvenile.treated.high,
                                   adult.control.ambient = adult.control.ambient,
                                   adult.control.high = adult.control.high,
                                   adult.treated.high = adult.treated.high)
```


# Functions
```{r function-delta-Cq}
calculate_delta_Cq <- function(df) {
  df <- df %>%
    group_by(Sample) %>%
    mutate(delta_Cq = Cq.Mean - Cq.Mean[Target == "GAPDH"]) %>%
    ungroup()
  
  return(df)
}
```

## SEED AND SPAT ANALYSIS
### Read in files
```{r read-in-files}
# Set the directory where your CSV files are located
cqs_directory <- "lifestage_carryover/data/qPCR/Cq"

# Get a list of all CSV files in the directory with the naming structure "*Cq_Results.csv"
cq_file_list <- list() # Initialize list
cq_file_list <- list.files(path = cqs_directory, pattern = "Cq_Results\\.csv$", full.names = TRUE)

# Initialize an empty list to store the data frames
data_frames <- list()

# Loop through each file and read it into a data frame, then add it to the list
for (file in cq_file_list) {
  data <- read.csv(file, header = TRUE)
  data_frames[[file]] <- data
}

# Combine all data frames into a single data frame
combined_df <- bind_rows(data_frames, .id = "data_frame_id")

# Convert Sample column to character type
combined_df <- combined_df %>%
  mutate(Sample = as.character(Sample))

str(combined_df)

```

### Unique samples by target
```{r unqique-samples-by-target}
# Group by Sample and Target, then summarize to get unique rows for each sample
aggregated_df <- combined_df %>%
  group_by(Sample, Target) %>%
  summarize(Cq.Mean = mean(Cq.Mean, na.rm = TRUE)) %>%
  ungroup()

str(aggregated_df)
```

### Add life stage and treatment cols
```{r add-cols}
# Initialize new columns
aggregated_df <- aggregated_df %>%
  mutate(life.stage = NA_character_,
         treatment = NA_character_)

# Loop through each vector
for (vec_name in names(seed.spat_vector_list)) {
  vec <- seed.spat_vector_list[[vec_name]]
  stage <- strsplit(vec_name, "\\.")[[1]][1]
  treatment <- strsplit(vec_name, "\\.")[[1]][2]
  
  # Loop through each row in aggregated_df
  for (i in 1:nrow(aggregated_df)) {
    sample <- aggregated_df$Sample[i]
    
    # Check if sample is in the vector
    if (sample %in% vec) {
      # Update life.stage and treatment columns
      aggregated_df$life.stage[i] <- stage
      aggregated_df$treatment[i] <- treatment
    }
  }
}

str(aggregated_df)

```

### Delta Cq to Normalizing Gene
```{r delta-Cq}
# Calculate delta Cq by subtracting GAPDH Cq.Mean from each corresponding Sample Cq.Mean
delta_Cq_df <- calculate_delta_Cq(aggregated_df)

str(delta_Cq_df)
```


### T-tests
```{r t-tests}
# Filter out groups with missing life.stage or Target
# Caused by NTCs
# Also removes normalizing gene(s)
delta_Cq_df_filtered <- delta_Cq_df %>%
  filter(!is.na(life.stage), !is.na(Target), Target != "GAPDH")

# Perform t-test for each Target within life.stage
t_test_results <- delta_Cq_df_filtered %>%
  group_by(life.stage, Target) %>%
  summarise(
    t_test_result = list(t.test(delta_Cq ~ treatment))
  ) %>%
  ungroup()

# Extract t-test statistics
t_test_results <- t_test_results %>%
  mutate(
    estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
    p_value = sapply(t_test_result, function(x) x$p.value)
  ) %>% 
  select(!t_test_result)

# Add asterisk information to data frame
# if p-value <= 0.05
# Useful for plotting
t_test_results$asterisk <- ifelse(t_test_results$p_value <= 0.05, "*", "")

str(t_test_results)
```

### Plotting

#### Delta Cq Box Plots

##### Seed

```{r box-plots-seed-delta-Cq}
library(ggplot2)

# Filter delta_Cq_df_filtered for seed life stage
seed_delta_Cq_df <- delta_Cq_df_filtered %>%
  filter(life.stage == "seed")

# Create the box plot
seed_boxplot <- ggplot(seed_delta_Cq_df, aes(x = Target, y = delta_Cq, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75)) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_fill_manual(values=c("darkgray", "salmon"))+
  ylim(-2,10)+
  labs(x = "Target", y = "Delta Cq", title = "Seed") + 
  annotate("text", x = t_test_results$Target, y = 8, label = t_test_results$asterisk, # Adds asterisks
           vjust = -0.5, size = 8, color = "black")

# Show box plot
print(seed_boxplot)
```

##### Spat
```{r box-plots-delta-Cq-spat}
# Filter data for life.stage = "spat"
spat_delta_Cq <- delta_Cq_df_filtered %>%
  filter(life.stage == "spat")

# Calculate the maximum delta_Cq for each Target
max_delta_Cq_by_target <- spat_delta_Cq %>%
  group_by(Target) %>%
  summarise(max_delta_Cq = max(delta_Cq, na.rm = TRUE))

# Merge t_test_results with max_delta_Cq_by_target to get the maximum delta_Cq for each Target
t_test_results_with_max_delta_Cq <- merge(t_test_results, max_delta_Cq_by_target, by = "Target")


# Create the box plot
spat_boxplot <- ggplot(spat_delta_Cq, aes(x = Target, y = delta_Cq, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75)) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_fill_manual(values=c("darkgray", "salmon"))+
  labs(x = "Target", y = "Delta Cq", title = "Spat") +
  ylim(-2, 10)+
  annotate("text", x = t_test_results_with_max_delta_Cq$Target, # Adds asterisks
           y = 8, 
           label = t_test_results_with_max_delta_Cq$asterisk,
           vjust = -0.5, size = 10, color = "black")

# Show box plot
print(spat_boxplot)
```

### Delta delta Cq

#### Add treatment and life stage

```{r add-treatment-lifestage}
# Initialize empty vectors to store life.stage and treatment
life_stage <- character(nrow(combined_df))
treatment <- character(nrow(combined_df))

# Loop through each row of combined_df
for (i in 1:nrow(combined_df)) {
  sample_id <- combined_df$Sample[i]
  
  # Check if the sample_id is present in any of the vectors
  found <- FALSE
  for (vec_name in names(seed.spat_vector_list)) {
    if (sample_id %in% seed.spat_vector_list[[vec_name]]) {
      # If present, extract life.stage and treatment from the vector name
      parts <- strsplit(vec_name, "\\.")[[1]]
      life_stage[i] <- parts[1]
      treatment[i] <- parts[2]
      found <- TRUE
      break  # Exit loop once found
    }
  }
  
  # If sample_id is not found in any vector, assign NA to both life.stage and treatment
  if (!found) {
    life_stage[i] <- NA
    treatment[i] <- NA
  }
}

# Add life.stage and treatment columns to combined_df
combined_df <- combined_df %>%
  mutate(life.stage = life_stage,
         treatment = treatment)

# Filter out rows where life.stage is NA
combined_df_filtered <- combined_df %>% 
  filter(!is.na(life.stage))

str(combined_df_filtered)
```


#### Mean Cqs per gene per treatment per life stage
```{r mean-Cqs}
# Group by life.stage, treatment, and Target, then calculate the mean Cq
mean_Cq_df <- combined_df_filtered %>%
  group_by(life.stage, treatment, Target) %>%
  summarise(mean_Cq = mean(Cq, na.rm = TRUE))

str(mean_Cq_df)
```

#### Delta Cqs

```{r delta-Cqs-gene-level}
combined_df_with_delta_Cq <- mean_Cq_df %>%
  group_by(life.stage, treatment) %>%
  mutate(delta_Cq = mean_Cq - mean_Cq[Target == "GAPDH"]) %>%
  ungroup() %>%
  filter(Target != "GAPDH")

str(combined_df_with_delta_Cq)
```

#### Delta delta Cq

```{r delta-delta-Cq}
# Calculate delta_delta_Cq
delta_delta_Cq_df <- combined_df_with_delta_Cq %>%
  group_by(life.stage, Target) %>%
  summarize(delta_delta_Cq = delta_Cq[treatment == "control"] - delta_Cq[treatment == "treated"])

str(delta_delta_Cq_df)

```

#### Calculate the fold change for each Target
```{r fold-change}
delta_delta_fold_change <- delta_delta_Cq_df %>%
  mutate(fold_change = 2^(-delta_delta_Cq)) %>% 
  distinct(Target, fold_change)

str(delta_delta_fold_change)
```

### Plot - Seed Fold Change

```{r plot-seed-fold-change}
# Filter delta_delta_fold_change for seed life stage
seed_df <- delta_delta_fold_change %>%
  filter(life.stage == "seed")

# Create bar plot for seed life stage
# Create the plot with fold changes relative to baseline of 1
seed_fc_plot <- ggplot(seed_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "white", color="black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth=2) +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Seed Life Stage Fold Change") +
  scale_y_continuous(limits = c(-0.5, 1))+
  
  geom_text(aes(x=2, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=5, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=6, y=0.65, label="*"), color="black", size=10)

# Display plot
seed_fc_plot

```

### Plot - Spat Fold Change
```{r plot-spat-fold-change}
# Filter delta_delta_fold_change for spat life stage
spat_df <- delta_delta_fold_change %>%
  filter(life.stage == "spat")

# Create bar plot for spat life stage
# Create the plot with fold changes relative to baseline of 1
spat_fc_plot <- ggplot(spat_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "white", color="black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth=2) +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Spat Life Stage Fold Change") +
  scale_y_continuous(limits = c(-1, 0.5))+
  
  geom_text(aes(x=2, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=5, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=6, y=0.2, label="*"), color="black", size=10)
  

# Display plot
spat_fc_plot
```

# Generate panel plots

Generate seed plot
```{r}
library(cowplot)

seed_grid<-plot_grid(seed_boxplot, seed_fc_plot, ncol=2, nrow=1, align="vh", rel_widths=c(1,0.8))
ggsave(plot=seed_grid, "lifestage_carryover/output/qPCR/seed_qPCR.jpeg", width=12, height=4)

spat_grid<-plot_grid(spat_boxplot, spat_fc_plot, ncol=2, nrow=1, align="vh", rel_widths=c(1,0.8))
ggsave(plot=spat_grid, "lifestage_carryover/output/qPCR/spat_qPCR.jpeg", width=12, height=4)

```

Generate spat plot

```

