---
title: "01.01-qPCR"
author: "Sam White"
date: "2024-03-27"
output: 
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  github_document:
    toc: true
    number_sections: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
bibliography: references.bib
link-citations: true
---

# Load libraries
```{r load-libraries}
library("tidyverse")
library("ggplot2")
```

# Set variables

## Set sample groups

Groups are named in the following fashion:

`<life.stage>.<conditioning.treatment>.<acute.treatment>`

This allows for parsing downstream.
```{r set-sample-groups}

seed.control.ambient=c("29", "40", "55", "63", "69", "101", "119", "122", "155", "164", "187", "202", "209", "214", "233", "236", "275")
seed.control.high=c("42", "59", "60", "62", "86", "102", "140", "176", "177", "184", "192", "223", "234", "243", "244", "254", "264")
seed.treated.ambient=c("14", "48", "66", "72", "89", "115", "129", "138", "156", "182", "191", "201", "227", "239", "270", "277", "280")
seed.treated.high=c("15", "19", "24", "88", "92", "105", "111", "113", "120", "128", "161", "200", "211", "256", "257", "266", "285")
spat.control.ambient=c("11", "30", "36", "52", "77", "114", "134", "142", "144", "183", "193", "229", "230", "231", "240", "272", "287")
spat.control.high=c("27", "74", "93", "96", "97", "137", "143", "153", "168", "178", "189", "206", "262", "274", "282", "284", "289")
spat.treated.ambient=c("9", "13", "38", "46", "47", "121", "145", "151", "174", "194", "197", "198", "216", "235", "241", "252", "291")
spat.treated.high=c("6", "25", "50", "78", "124", "126", "131", "160", "163", "172", "220", "226", "242", "253", "296", "298")
juvenile.control.ambient=c("18", "57", "65", "75", "79", "104", "110", "123", "125", "171", "175", "205", "238", "273", "279", "293", "317")
juvenile.control.high=c("12", "39", "43", "49", "71", "130", "141", "146", "150", "170", "195", "297", "301", "324", "351", "355", "371")
juvenile.treated.ambient=c("1", "34", "64", "83", "98", "147", "152", "158", "162", "169", "188", "271", "295", "310", "357", "361", "381")
juvenile.treated.high=c("28", "53", "61", "73", "81", "106", "109", "139", "149", "173", "181", "213", "290", "302", "311", "364", "392")
adult.control.ambient=c("3", "5", "13*", "16", "17", "80", "87", "94", "148", "159", "179", "180", "250", "258", "268", "312", "326", "330", "334", "346", "360", "377", "379", "386")
adult.control.high=c("20", "23", "26", "32", "33", "67", "70", "90", "107", "132", "135", "157", "166", "186", "207", "215", "248", "316", "341", "344", "349", "382", "394", "395")
adult.treated.ambient=c("7", "31", "35", "37", "41", "54", "84", "100", "112", "116", "118", "133", "154", "199", "203", "204", "208", "219", "294", "318", "339", "353", "363", "378")
adult.treated.high=c("21", "22", "45", "82", "85", "91", "95", "99", "103", "108", "117", "127", "165", "185", "190", "196", "232", "237", "245", "263", "276", "306", "343", "374")

```

```{r assign-groups-to-list}
# Combine vectors into lists
# Used for adding treatment info and/or subsetting downstream

groups_list <- list(juvenile.control.ambient = juvenile.control.ambient,
                                   juvenile.control.high = juvenile.control.high,
                                   juvenile.treated.ambient = juvenile.treated.ambient,
                                   juvenile.treated.high = juvenile.treated.high,
                                   adult.control.ambient = adult.control.ambient,
                                   adult.control.high = adult.control.high,
                                   adult.treated.ambient = adult.treated.ambient,
                                   adult.treated.high = adult.treated.high,
                                   seed.control.ambient = seed.control.ambient,
                                   seed.control.high = seed.control.high,
                                   seed.treated.ambient = seed.treated.ambient,
                                   seed.treated.high = seed.treated.high,
                                   spat.control.ambient = spat.control.ambient,
                                   spat.control.high = spat.control.high,
                                   spat.treated.ambient = spat.treated.ambient,
                                   spat.treated.high = spat.treated.high)
```

# Functions

## Calculate delta Cq
```{r function-delta-Cq}
calculate_delta_Cq <- function(df) {
  df <- df %>%
    group_by(Sample) %>%
    mutate(delta_Cq = Cq.Mean - Cq.Mean[Target == "GAPDH"]) %>%
    ungroup()
  
  return(df)
}
```

## Create delta Cq boxplots

### Lifestage comparison
```{r function-boxplots-delta-Cq}
# Function to create box plots for each comparison
create_boxplot_delta_Cq <- function(data, comparison, t_test_results) {
  # Extract life stages from comparison
  life_stages <- unlist(strsplit(comparison, "\\."))

  # Debugging: Print life stages
  print(paste("Life stages for comparison:", comparison))
  print(life_stages)

  # Filter data for the relevant life stages
  filtered_data <- data %>%
    filter(life.stage %in% life_stages)

  # Debugging: Print filtered data
  print("Filtered data:")
  print(filtered_data)

  # Check if both life stages are included
  if (!all(life_stages %in% unique(filtered_data$life.stage))) {
    stop("Not all life stages are included in the filtered data")
  }

  y_limits <- range(filtered_data$delta_Cq, na.rm = TRUE)

  # Debugging: Print y_limits
  print("Y limits:")
  print(y_limits)

  # Filter t_test_results for the current comparison
  t_test_results_filtered <- t_test_results %>%
    filter(comparison == !!comparison)

  # Debugging: Print filtered t_test_results
  print("Filtered t_test_results:")
  print(t_test_results_filtered)

  # Filter t_test_results for asterisks
  t_test_results_with_asterisks <- t_test_results_filtered %>%
    filter(asterisk != "")

  # Debugging: Print t_test_results_with_asterisks
  print("t_test_results_with_asterisks:")
  print(t_test_results_with_asterisks)

  formatted_title <- paste0(toupper(substring(life_stages[1], 1, 1)), substring(life_stages[1], 2), 
                            " vs. ", 
                            toupper(substring(life_stages[2], 1, 1)), substring(life_stages[2], 2))

  boxplot <- ggplot(filtered_data, aes(x = Target, y = delta_Cq, fill = life.stage)) +
    geom_boxplot(position = position_dodge(width = 0.75)) +
    theme_minimal() +
    theme(legend.position = "right") +
    scale_fill_manual(values=c("darkgray", "salmon", "lightblue", "lightgreen")) +
    ylim(y_limits) +
    labs(x = "Target", y = "Delta Cq", title = formatted_title) +
    # Highlighted section: Adds asterisks
    geom_text(data = t_test_results_with_asterisks, 
              aes(x = Target, y = y_limits[2] - 1, label = asterisk), 
              vjust = -0.5, size = 8, color = "black", inherit.aes = FALSE)

  print(boxplot)
}
```

### Treatment comparisons
```{r function-boxplots-delta-Cq-treatments}
# Function to create box plots for each comparison of conditioning treatments within life stages
create_boxplot_conditioning <- function(data, comparison, t_test_results) {
  # Extract life stage and conditioning treatments from comparison
  comparison_parts <- unlist(strsplit(comparison, "\\."))
  life_stage <- comparison_parts[1]
  treatment1 <- comparison_parts[2]
  treatment2 <- comparison_parts[3]

  # Debugging: Print life stage and treatments
  print(paste("Life stage and treatments for comparison:", comparison))
  print(c(life_stage, treatment1, treatment2))

  # Filter data for the relevant life stage and conditioning treatments
  filtered_data <- data %>%
    filter(life.stage == life_stage, conditioning.treatment %in% c(treatment1, treatment2))

  # Debugging: Print filtered data
  print("Filtered data:")
  print(filtered_data)

  # Check if both treatments are included
  if (!all(c(treatment1, treatment2) %in% unique(filtered_data$conditioning.treatment))) {
    stop("Not all treatments are included in the filtered data")
  }

  y_limits <- range(filtered_data$delta_Cq, na.rm = TRUE)

  # Debugging: Print y_limits
  print("Y limits:")
  print(y_limits)

  # Filter t_test_results for the current comparison
  t_test_results_filtered <- t_test_results %>%
    filter(comparison == !!comparison)

  # Debugging: Print filtered t_test_results
  print("Filtered t_test_results:")
  print(t_test_results_filtered)

  # Filter t_test_results for asterisks
  t_test_results_with_asterisks <- t_test_results_filtered %>%
    filter(asterisk != "")

  # Debugging: Print t_test_results_with_asterisks
  print("t_test_results_with_asterisks:")
  print(t_test_results_with_asterisks)

  # Format the title
  formatted_title <- paste0(toupper(substring(life_stage, 1, 1)), substring(life_stage, 2), 
                            " - ", 
                            toupper(substring(treatment1, 1, 1)), substring(treatment1, 2), 
                            " vs. ", 
                            toupper(substring(treatment2, 1, 1)), substring(treatment2, 2))

  boxplot <- ggplot(filtered_data, aes(x = Target, y = delta_Cq, fill = conditioning.treatment)) +
    geom_boxplot(position = position_dodge(width = 0.75)) +
    theme_minimal() +
    theme(legend.position = "right") +
    scale_fill_manual(values=c("darkgray", "salmon")) +
    ylim(y_limits) +
    labs(x = "Target", y = "Delta Cq", title = formatted_title) +
    # Highlighted section: Adds asterisks
    geom_text(data = t_test_results_with_asterisks, 
              aes(x = Target, y = y_limits[2] - 1, label = asterisk), 
              vjust = -0.5, size = 8, color = "black", inherit.aes = FALSE)

  print(boxplot)
}
```

# Read in files
```{r read-in-Cq-files}
# Set the directory where your CSV files are located
cqs_directory <- "lifestage_carryover/data/qPCR/Cq"

# Get a list of all CSV files in the directory with the naming structure "*Cq-Results.csv"
cq_file_list <- list() # Initialize list
cq_file_list <- list.files(path = cqs_directory, pattern = "Cq-Results\\.csv$", full.names = TRUE)

# Initialize an empty list to store the data frames
data_frames_list <- list()

# Loop through each file and read it into a data frame, then add it to the list
for (file in cq_file_list) {
  data <- read.csv(file, header = TRUE)
  data$Sample <- as.character(data$Sample)  # Convert Sample column to character type
  data_frames_list[[file]] <- data
}

# Combine all data frames into a single data frame
combined_df <- bind_rows(data_frames_list, .id = "data_frame_id")

str(combined_df)

```

# Clean data


## Replace target names
```{r replace-target-names}
# Remove rows with Sample name "NTC"
combined_df <- combined_df[combined_df$Sample != "NTC", ]


# Replace values in the Target column
combined_df$Target <- gsub("Cg_GAPDH_205_F-355_R \\(SR IDs: 1172/3\\)", "GAPDH", combined_df$Target)

combined_df$Target <- gsub("Cg_ATPsynthase_F/R \\(SR IDs: 1385/6\\)", "ATPsynthase", combined_df$Target)

combined_df$Target <- gsub("Cg_cGAS \\(SR IDs: 1826/7\\)", "cGAS", combined_df$Target)

combined_df$Target <- gsub("Cg_citrate_synthase \\(SR IDs: 1383/4\\)", "citrate.sythase", combined_df$Target)

combined_df$Target <- gsub("Cg_DNMT1_F \\(SR IDs: 1510/1\\)", "DNMT1", combined_df$Target)

combined_df$Target <- gsub("Cg_HSP70_F/R \\(SR IDs: 598/9\\)", "HSP70", combined_df$Target)

combined_df$Target <- gsub("Cg_Hsp90_F/R \\(SR IDs: 1532/3\\)", "HSP90", combined_df$Target)

combined_df$Target <- gsub("Cg_VIPERIN_F/R \\(SR IDs: 1828/9\\)", "VIPERIN", combined_df$Target)

str(combined_df)
```

## Identify Samples with Cq.Std..Dev > 0.5
```{r high-cq-std-dev}
# Filter out rows where Cq.Std..Dev is NA
combined_df <- combined_df[!is.na(combined_df$Cq.Std..Dev), ]

# Filter rows where Cq.Std..Dev is greater than 0.5
high_cq_std_dev <- combined_df[combined_df$Cq.Std..Dev > 0.5, ]

# Print the filtered rows with specified columns, without row names
print(high_cq_std_dev[, c("Target", "Sample", "Cq", "Cq.Std..Dev")], row.names = FALSE)
```

## Remove bad technical reps
```{r adult-juv-remove-bad-technical-reps}
# Group by Sample and Target, then filter out the outlier replicate
combined.fitered_df<- combined_df %>%
  group_by(Sample, Target) %>%
  filter(abs(Cq - mean(Cq, na.rm = TRUE)) <= Cq.Std..Dev)

# Print the filtered data frame
str(combined.fitered_df)
```



# Group samples by target
```{r samples-by-target}
# Group by Sample and Target, then summarize to get unique rows for each sample
grouped_df <- combined.fitered_df%>%
  group_by(Sample, Target) %>%
  summarize(Cq.Mean = mean(Cq, na.rm = TRUE)) %>%
  ungroup()

str(grouped_df)
```

# Add life stage and treatment cols
```{r add-columns}
# Initialize new columns
grouped_df <- grouped_df %>%
  mutate(life.stage = NA_character_,
         conditioning.treatment = NA_character_,
         acute.treatment = NA_character_)

# Loop through each vector
for (vec_name in names(groups_list)) {
  vec <- groups_list[[vec_name]]
  stage <- strsplit(vec_name, "\\.")[[1]][1]
  conditioning_treatment <- strsplit(vec_name, "\\.")[[1]][2]
  acute_treatment <- strsplit(vec_name, "\\.")[[1]][3]
  
  # Loop through each row in grouped_df
  for (i in 1:nrow(grouped_df)) {
    sample <- grouped_df$Sample[i]
    
    # Check if sample is in the vector
    if (sample %in% vec) {
      # Update life.stage and treatment columns
      grouped_df$life.stage[i] <- stage
      grouped_df$conditioning.treatment[i] <- conditioning_treatment
      grouped_df$acute.treatment[i] <-acute_treatment
    }
  }
}

str(grouped_df)

```

# Delta Cq to Normalizing Gene
```{r delta-Cq}
# Calculate delta Cq by subtracting GAPDH Cq.Mean from each corresponding Sample Cq.Mean
delta_Cq_df <- calculate_delta_Cq(grouped_df)

delta_Cq_df <- delta_Cq_df %>%
  filter(!is.na(life.stage), !is.na(Target), Target != "GAPDH")

str(delta_Cq_df)
```

# t-tests

## Life Stages

This code does the following:

1. Extracts the unique life.stage levels from the data frame.
2. Generates all possible pairs of life.stage levels using the combn function.
3. Iterates over each pair and performs the t-test for each Target. Adds an
asterisk column and an asterisk if the p-value is <= 0.05. Useful for downstream
parsing.
4. Stores the results in a list and combines them into a single data frame.
5. Adds a comparison column to indicate which life.stage levels were compared.

```{r life-stage-t-tests}
# Extract unique life.stage levels
unique_life_stages <- unique(delta_Cq_df$life.stage)

# Generate all possible pairs of life.stage levels
life_stage_pairs <- combn(unique_life_stages, 2, simplify = FALSE)

# Initialize a list to store results
life_stage_t_test_results_list <- list()

for (pair in life_stage_pairs) {
  stage1 <- pair[1]
  stage2 <- pair[2]
  
  # Perform t-test for each Target comparing the two life.stage levels
  t_test_results <- delta_Cq_df %>%
    filter(life.stage %in% c(stage1, stage2)) %>%
    group_by(Target) %>%
    summarise(
      t_test_result = list(t.test(delta_Cq ~ life.stage))
    ) %>%
    ungroup() %>%
    mutate(
      estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
      p_value = sapply(t_test_result, function(x) x$p.value),
      asterisk = ifelse(p_value <= 0.05, "*", ""), # Adds asterisk column and asterisk for p-value.
      comparison = paste(stage1, "vs", stage2, sep = ".")
    ) %>%
    select(!t_test_result)
  
  life_stage_t_test_results_list[[paste(stage1, stage2, sep = ".")]] <- t_test_results
}

# Combine results into a single data frame
life_stage_t_test_results_df <- bind_rows(life_stage_t_test_results_list, .id = "comparison")

# View the results
print(life_stage_t_test_results_df)
```

## Conditioning treatments

This code does the following:

1. Extracts the unique life.stage levels from the data frame.
2. For each life.stage, extracts the unique conditioning.treatment levels.
3. Generates all possible pairs of conditioning.treatment levels within each life.stage.
4. Iterates over each pair and performs the t-test for each Target. Adds an
asterisk column and an asterisk if the p-value is <= 0.05. Useful for downstream
parsing.
5. Stores the results in a list and combines them into a single data frame.
6. Adds a comparison column to indicate which life.stage and conditioning.treatment levels were compared.

```{r conditioning-treatments-life-stages}
# Extract unique life.stage levels
unique_life_stages <- unique(delta_Cq_df$life.stage)

# Initialize a list to store results
conditioning_treatment_t_test_results_list <- list()

for (stage in unique_life_stages) {
  # Extract unique conditioning.treatment levels within the current life.stage
  unique_treatments <- unique(delta_Cq_df %>% filter(life.stage == stage) %>% pull(conditioning.treatment))
  
  # Generate all possible pairs of conditioning.treatment levels
  treatment_pairs <- combn(unique_treatments, 2, simplify = FALSE)
  
  for (pair in treatment_pairs) {
    treatment1 <- pair[1]
    treatment2 <- pair[2]
    
    # Perform t-test for each Target comparing the two conditioning.treatment levels within the current life.stage
    t_test_results <- delta_Cq_df %>%
      filter(life.stage == stage, conditioning.treatment %in% c(treatment1, treatment2)) %>%
      group_by(Target) %>%
      summarise(
        t_test_result = list(t.test(delta_Cq ~ conditioning.treatment))
      ) %>%
      ungroup() %>%
      mutate(
        estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
        p_value = sapply(t_test_result, function(x) x$p.value),
        asterisk = ifelse(p_value <= 0.05, "*", ""), # Adds asterisk column and asterisk for p-value.
        comparison = paste(stage, treatment1, "vs", treatment2, sep = ".")
      ) %>%
      select(!t_test_result)
    
    conditioning_treatment_t_test_results_list[[paste(stage, treatment1, treatment2, sep = ".")]] <- t_test_results
  }
}

# Combine results into a single data frame
conditioning_treatment_t_test_results_df <- bind_rows(conditioning_treatment_t_test_results_list, .id = "comparison")

# View the results
print(conditioning_treatment_t_test_results_df)
```


## Acute treatments

This code does the following:

1. Extracts the unique life.stage levels from the data frame.
2. For each life.stage, extracts the unique acute.treatment levels.
3. Generates all possible pairs of acute.treatment levels within each life.stage.
4. Iterates over each pair and performs the t-test for each Target. Adds an
asterisk column and an asterisk if the p-value is <= 0.05. Useful for downstream
parsing.
5. Stores the results in a list and combines them into a single data frame.
6. Adds a comparison column to indicate which life.stage and acute.treatment levels were compared.

Excludes `seed` and `spat`, as these were only held at `ambient` for the acute treatment.
```{r acute-treatments-life-stages}
# Extract unique life.stage levels, excluding 'seed' and 'spat'
unique_life_stages <- unique(delta_Cq_df$life.stage)
unique_life_stages <- setdiff(unique_life_stages, c("seed", "spat"))

# Initialize a list to store results
acute_treatment_t_test_results_list <- list()

for (stage in unique_life_stages) {
  # Extract unique acute.treatment levels within the current life.stage
  unique_treatments <- unique(delta_Cq_df %>% filter(life.stage == stage) %>% pull(acute.treatment))
  
  # Check if there are at least 2 unique treatments
  if (length(unique_treatments) >= 2) {
    # Generate all possible pairs of acute.treatment levels
    treatment_pairs <- combn(unique_treatments, 2, simplify = FALSE)
    
    for (pair in treatment_pairs) {
      treatment1 <- pair[1]
      treatment2 <- pair[2]
      
      # Perform t-test for each Target comparing the two acute.treatment levels within the current life.stage
      t_test_results <- delta_Cq_df %>%
        filter(life.stage == stage, acute.treatment %in% c(treatment1, treatment2)) %>%
        group_by(Target) %>%
        summarise(
          t_test_result = list(t.test(delta_Cq ~ acute.treatment))
        ) %>%
        ungroup() %>%
        mutate(
          estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
          p_value = sapply(t_test_result, function(x) x$p.value),
          asterisk = ifelse(p_value <= 0.05, "*", ""), # Adds asterisk column and asterisk for p-value.
          comparison = paste(stage, treatment1, "vs", treatment2, sep = ".")
        ) %>%
        select(!t_test_result)
      
      acute_treatment_t_test_results_list[[paste(stage, treatment1, treatment2, sep = ".")]] <- t_test_results
    }
  }
}

# Combine results into a single data frame
acute_treatment_t_test_results_df <- bind_rows(acute_treatment_t_test_results_list, .id = "comparison")

# View the results
print(acute_treatment_t_test_results_df)

```

# Plotting

## Delta Cq boxplots

## Lifestage comparisons
```{r delta-Cq-boxplots-lifestage}
# Create box plots for each comparison
unique_comparisons <- unique(life_stage_t_test_results_df$comparison)

for (comparison in unique_comparisons) {
  create_boxplot_delta_Cq(delta_Cq_df, comparison, life_stage_t_test_results_df)
}
```

## Conditioning comparisons
```{r delta-Cq-boxplots-conditioning}

# Create box plots for each comparison
unique_comparisons <- unique(conditioning_treatment_t_test_results_df$comparison)

for (comparison in unique_comparisons) {
  create_boxplot_conditioning(delta_Cq_df, comparison, conditioning_treatment_t_test_results_df)
}
```

---

# PRELIMINARY ANALYSIS

Analysis below was previously conducted with a preliminary set of samples to decide
if this project warranted moving forward with.

## Set sample groups
```{r set-sample-groups}
seed.control=c("223", "243", "244")
seed.treated=c("200", "257", "285")
spat.control=c("206", "282", "284", "289")
spat.treated=c("220", "226", "242", "253", "296", "298")

# Combine vectors into lists
# Used for adding treatment info and/or subsetting downstream
seed.spat_vector_list <- list(seed.control = seed.control,
                    seed.treated = seed.treated,
                    spat.control = spat.control,
                    spat.treated = spat.treated)

```

### Read in files
```{r read-in-files}
# Set the directory where your CSV files are located
cqs_directory <- "lifestage_carryover/data/qPCR/Cq"

# Get a list of all CSV files in the directory with the naming structure "*Cq_Results.csv"
cq_file_list <- list() # Initialize list
cq_file_list <- list.files(path = cqs_directory, pattern = "Cq_Results\\.csv$", full.names = TRUE)

# Initialize an empty list to store the data frames
data_frames <- list()

# Loop through each file and read it into a data frame, then add it to the list
for (file in cq_file_list) {
  data <- read.csv(file, header = TRUE)
  data_frames[[file]] <- data
}

# Combine all data frames into a single data frame
combined_df <- bind_rows(data_frames, .id = "data_frame_id")

# Convert Sample column to character type
combined_df <- combined_df %>%
  mutate(Sample = as.character(Sample))

str(combined_df)

```

### Unique samples by target
```{r unqique-samples-by-target}
# Group by Sample and Target, then summarize to get unique rows for each sample
aggregated_df <- combined_df %>%
  group_by(Sample, Target) %>%
  summarize(Cq.Mean = mean(Cq.Mean, na.rm = TRUE)) %>%
  ungroup()

str(aggregated_df)
```

### Add life stage and treatment cols
```{r add-cols}
# Initialize new columns
aggregated_df <- aggregated_df %>%
  mutate(life.stage = NA_character_,
         treatment = NA_character_)

# Loop through each vector
for (vec_name in names(seed.spat_vector_list)) {
  vec <- seed.spat_vector_list[[vec_name]]
  stage <- strsplit(vec_name, "\\.")[[1]][1]
  treatment <- strsplit(vec_name, "\\.")[[1]][2]
  
  # Loop through each row in aggregated_df
  for (i in 1:nrow(aggregated_df)) {
    sample <- aggregated_df$Sample[i]
    
    # Check if sample is in the vector
    if (sample %in% vec) {
      # Update life.stage and treatment columns
      aggregated_df$life.stage[i] <- stage
      aggregated_df$treatment[i] <- treatment
    }
  }
}

str(aggregated_df)

```

### Delta Cq to Normalizing Gene
```{r delta-Cq}
# Calculate delta Cq by subtracting GAPDH Cq.Mean from each corresponding Sample Cq.Mean
delta_Cq_df <- calculate_delta_Cq(aggregated_df)

str(delta_Cq_df)
```


### T-tests
```{r t-tests}
# Filter out groups with missing life.stage or Target
# Caused by NTCs
# Also removes normalizing gene(s)
delta_Cq_df_filtered <- delta_Cq_df %>%
  filter(!is.na(life.stage), !is.na(Target), Target != "GAPDH")

# Perform t-test for each Target within life.stage
t_test_results <- delta_Cq_df_filtered %>%
  group_by(life.stage, Target) %>%
  summarise(
    t_test_result = list(t.test(delta_Cq ~ treatment))
  ) %>%
  ungroup()

# Extract t-test statistics
t_test_results <- t_test_results %>%
  mutate(
    estimate_diff = sapply(t_test_result, function(x) x$estimate[1] - x$estimate[2]),
    p_value = sapply(t_test_result, function(x) x$p.value)
  ) %>% 
  select(!t_test_result)

# Add asterisk information to data frame
# if p-value <= 0.05
# Useful for plotting
t_test_results$asterisk <- ifelse(t_test_results$p_value <= 0.05, "*", "")

str(t_test_results)
```

### Plotting

#### Delta Cq Box Plots

##### Seed

```{r box-plots-seed-delta-Cq}
library(ggplot2)

# Filter delta_Cq_df_filtered for seed life stage
seed_delta_Cq_df <- delta_Cq_df_filtered %>%
  filter(life.stage == "seed")

# Create the box plot
seed_boxplot <- ggplot(seed_delta_Cq_df, aes(x = Target, y = delta_Cq, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75)) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_fill_manual(values=c("darkgray", "salmon"))+
  ylim(-2,10)+
  labs(x = "Target", y = "Delta Cq", title = "Seed") + 
  annotate("text", x = t_test_results$Target, y = 8, label = t_test_results$asterisk, # Adds asterisks
           vjust = -0.5, size = 8, color = "black")

# Show box plot
print(seed_boxplot)
```

##### Spat
```{r box-plots-delta-Cq-spat}
# Filter data for life.stage = "spat"
spat_delta_Cq <- delta_Cq_df_filtered %>%
  filter(life.stage == "spat")

# Calculate the maximum delta_Cq for each Target
max_delta_Cq_by_target <- spat_delta_Cq %>%
  group_by(Target) %>%
  summarise(max_delta_Cq = max(delta_Cq, na.rm = TRUE))

# Merge t_test_results with max_delta_Cq_by_target to get the maximum delta_Cq for each Target
t_test_results_with_max_delta_Cq <- merge(t_test_results, max_delta_Cq_by_target, by = "Target")


# Create the box plot
spat_boxplot <- ggplot(spat_delta_Cq, aes(x = Target, y = delta_Cq, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75)) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_fill_manual(values=c("darkgray", "salmon"))+
  labs(x = "Target", y = "Delta Cq", title = "Spat") +
  ylim(-2, 10)+
  annotate("text", x = t_test_results_with_max_delta_Cq$Target, # Adds asterisks
           y = 8, 
           label = t_test_results_with_max_delta_Cq$asterisk,
           vjust = -0.5, size = 10, color = "black")

# Show box plot
print(spat_boxplot)
```

### Delta delta Cq

#### Add treatment and life stage

```{r add-treatment-lifestage}
# Initialize empty vectors to store life.stage and treatment
life_stage <- character(nrow(combined_df))
treatment <- character(nrow(combined_df))

# Loop through each row of combined_df
for (i in 1:nrow(combined_df)) {
  sample_id <- combined_df$Sample[i]
  
  # Check if the sample_id is present in any of the vectors
  found <- FALSE
  for (vec_name in names(seed.spat_vector_list)) {
    if (sample_id %in% seed.spat_vector_list[[vec_name]]) {
      # If present, extract life.stage and treatment from the vector name
      parts <- strsplit(vec_name, "\\.")[[1]]
      life_stage[i] <- parts[1]
      treatment[i] <- parts[2]
      found <- TRUE
      break  # Exit loop once found
    }
  }
  
  # If sample_id is not found in any vector, assign NA to both life.stage and treatment
  if (!found) {
    life_stage[i] <- NA
    treatment[i] <- NA
  }
}

# Add life.stage and treatment columns to combined_df
combined_df <- combined_df %>%
  mutate(life.stage = life_stage,
         treatment = treatment)

# Filter out rows where life.stage is NA
combined_df_filtered <- combined_df %>% 
  filter(!is.na(life.stage))

str(combined_df_filtered)
```


#### Mean Cqs per gene per treatment per life stage
```{r mean-Cqs}
# Group by life.stage, treatment, and Target, then calculate the mean Cq
mean_Cq_df <- combined_df_filtered %>%
  group_by(life.stage, treatment, Target) %>%
  summarise(mean_Cq = mean(Cq, na.rm = TRUE))

str(mean_Cq_df)
```

#### Delta Cqs

```{r delta-Cqs-gene-level}
combined_df_with_delta_Cq <- mean_Cq_df %>%
  group_by(life.stage, treatment) %>%
  mutate(delta_Cq = mean_Cq - mean_Cq[Target == "GAPDH"]) %>%
  ungroup() %>%
  filter(Target != "GAPDH")

str(combined_df_with_delta_Cq)
```

#### Delta delta Cq

```{r delta-delta-Cq}
# Calculate delta_delta_Cq
delta_delta_Cq_df <- combined_df_with_delta_Cq %>%
  group_by(life.stage, Target) %>%
  summarize(delta_delta_Cq = delta_Cq[treatment == "control"] - delta_Cq[treatment == "treated"])

str(delta_delta_Cq_df)

```

#### Calculate the fold change for each Target
```{r fold-change}
delta_delta_fold_change <- delta_delta_Cq_df %>%
  mutate(fold_change = 2^(-delta_delta_Cq)) %>% 
  distinct(Target, fold_change)

str(delta_delta_fold_change)
```

### Plot - Seed Fold Change

```{r plot-seed-fold-change}
# Filter delta_delta_fold_change for seed life stage
seed_df <- delta_delta_fold_change %>%
  filter(life.stage == "seed")

# Create bar plot for seed life stage
# Create the plot with fold changes relative to baseline of 1
seed_fc_plot <- ggplot(seed_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "white", color="black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth=2) +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Seed Life Stage Fold Change") +
  scale_y_continuous(limits = c(-0.5, 1))+
  
  geom_text(aes(x=2, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=5, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=6, y=0.65, label="*"), color="black", size=10)

# Display plot
seed_fc_plot

```

### Plot - Spat Fold Change
```{r plot-spat-fold-change}
# Filter delta_delta_fold_change for spat life stage
spat_df <- delta_delta_fold_change %>%
  filter(life.stage == "spat")

# Create bar plot for spat life stage
# Create the plot with fold changes relative to baseline of 1
spat_fc_plot <- ggplot(spat_df, aes(x = Target, y = fold_change - 1)) +
  geom_bar(stat = "identity", fill = "white", color="black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth=2) +  # Baseline
  theme_minimal() +
  labs(x = "Target", y = "Fold Change", title = "Spat Life Stage Fold Change") +
  scale_y_continuous(limits = c(-1, 0.5))+
  
  geom_text(aes(x=2, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=5, y=0.2, label="*"), color="black", size=10)+
  geom_text(aes(x=6, y=0.2, label="*"), color="black", size=10)
  

# Display plot
spat_fc_plot
```

# Generate panel plots

Generate seed plot
```{r}
library(cowplot)

seed_grid<-plot_grid(seed_boxplot, seed_fc_plot, ncol=2, nrow=1, align="vh", rel_widths=c(1,0.8))
ggsave(plot=seed_grid, "lifestage_carryover/output/qPCR/seed_qPCR.jpeg", width=12, height=4)

spat_grid<-plot_grid(spat_boxplot, spat_fc_plot, ncol=2, nrow=1, align="vh", rel_widths=c(1,0.8))
ggsave(plot=spat_grid, "lifestage_carryover/output/qPCR/spat_qPCR.jpeg", width=12, height=4)

```

Generate spat plot
