---
title: "C. gigas growth - lifestage carryover project" 
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load libraries 

```{r}
library(tidyverse)
library(ggplot2)
```

# Growth during Lifestage CarryOver project first phase (October - February)

`gigas-lengths-lco.csv`

## Load data 

Load in data frames. Adult, juvenile, seed, and spat data are in the lenghts-lco data frame for this phase.   

```{r}
growth<-read_csv("lifestage_carryover/data/size/gigas-lengths-lco.csv")

growth<-growth%>%
  select(tag, `life-stage`, `init-treat`, month, length.mm)%>%
  rename(lifestage=`life-stage`, treatment=`init-treat`)

growth$month<-factor(growth$month, levels=c("Oct", "Nov", "Jan", "Feb", "March", "April", "May"))
```

## Look at data for whole population together  

### Plot growth over months for each life stage and treatment 

Plot as box plot. Filter out adults that were sampled in November, since these were larger and are not present in later dates. 
```{r}
plot<-growth%>%
  filter(!tag=="Cat")%>%
  filter(!tag=="Dog")%>%
  filter(!tag=="Yellow")%>%
  filter(!tag=="Toad")%>%
  
  ggplot(aes(color=treatment, x=month, y=length.mm))+
  geom_boxplot(fill=NA, outlier.shape=NA)+
  geom_point(position=position_dodge(0.7), size=1)+
  facet_grid(~lifestage, scales="free_x")+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot

ggsave(plot=plot, "lifestage_carryover/output/size/mean_lengths_box.png", width=7, height=4)
```

Plot as mean +- standard error for all.  
```{r}
plot2<-growth%>%
  filter(!tag=="Cat")%>%
  filter(!tag=="Dog")%>%
  filter(!tag=="Yellow")%>%
  filter(!tag=="Toad")%>%
  
  group_by(treatment, month, lifestage)%>%
  summarise(mean=mean(length.mm, na.rm=TRUE), se=sd(length.mm, na.rm=TRUE)/sqrt(length(length.mm)))%>%
  
  ggplot(aes(color=treatment, x=month, y=mean))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=1)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, position=position_dodge(0.3), linewidth=0.75)+
  geom_line(aes(group=treatment), position=position_dodge(0.3), linewidth=1)+
  facet_grid(~lifestage, scales="free_x")+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot2

ggsave(plot=plot2, "lifestage_carryover/output/size/mean_lengths.png", width=7, height=4)
```

### Run model

```{r}
model<-aov(length.mm~lifestage*month*treatment, data=growth)
summary(model)
```

                            Df  Sum Sq Mean Sq  F value  Pr(>F)    
lifestage                    3 1187777  395926 9758.267 < 2e-16 ***
month                        6   12889    2148   52.943 < 2e-16 ***
treatment                    1     162     162    3.992 0.04580 *  
lifestage:month              6   21932    3655   90.094 < 2e-16 ***
lifestage:treatment          3     573     191    4.711 0.00277 ** 
month:treatment              6     103      17    0.423 0.86429    
lifestage:month:treatment    6     203      34    0.833 0.54413    
Residuals                 3282  133162      41         

Conduct posthoc test
```{r}
library(emmeans)
emm<-emmeans(model, ~ treatment | month | lifestage)
pairs(emm)
```

Size is different in October and November for adults. No other differences.  

Conduct model for all time points for only spat.  

```{r}
model<-growth%>%
  filter(lifestage=="spat")%>%
  
  aov(length.mm~month*treatment, data=.)

summary(model)
```

                  Df Sum Sq Mean Sq F value   Pr(>F)    
month              5    371   74.18   7.939 2.03e-07 ***
treatment          1     18   17.79   1.904   0.1678    
month:treatment    5    104   20.85   2.232   0.0487 *  
Residuals       2024  18910    9.34 

Conduct post hoc tests. 

```{r}
emm<-emmeans(model, ~ treatment | month)
pairs(emm)
```

Spat were smaller in treated condiction in Feb and March, but no difference in other months. 

## Plot size of only adults 

Plot boxplot 
```{r}
plot2a<-growth%>%
  filter(tag %in% c("Blue", "Pink", "Bird", "Orange"))%>%
  
  group_by(treatment, month, tag)%>%
  
  ggplot(aes(color=treatment, x=month, y=length.mm))+
  geom_point(aes(group=interaction(treatment, tag)), position=position_dodge(0.3), size=1)+
  geom_boxplot(outlier.shape=NA)+
  facet_grid(~tag)+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot2a

ggsave(plot=plot2a, "lifestage_carryover/output/size/boxplot_lengths_adults.png", width=7, height=4)
```

Plot as mean +- standard error for all.  
```{r}
plot2b<-growth%>%
  filter(tag %in% c("Blue", "Pink", "Bird", "Orange"))%>%
  
  group_by(treatment, month, tag)%>%
  summarise(mean=mean(length.mm, na.rm=TRUE), se=sd(length.mm, na.rm=TRUE)/sqrt(length(length.mm)))%>%
  
  ggplot(aes(color=treatment, x=month, y=mean))+
  geom_point(aes(group=interaction(treatment, tag)), position=position_dodge(0.3), size=1)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, position=position_dodge(0.3), linewidth=0.75)+
  geom_line(aes(group=treatment), position=position_dodge(0.3), linewidth=1)+
  facet_grid(~tag)+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot2b

ggsave(plot=plot2b, "lifestage_carryover/output/size/mean_lengths_adults.png", width=7, height=4)
```

## Look at growth as percent growth summarised at the level of replicate containers or tags 

Calculate mean size at each month for each replicate and calculate % growth (change/original).  

```{r}
replicates<-growth%>%
  group_by(treatment, month, lifestage, tag)%>%
  summarise(mean=mean(length.mm, na.rm=TRUE))%>%
  pivot_wider(names_from=month, values_from=mean)%>%
  mutate(perc_growth_Nov=((Nov-Oct)/Oct)*100)%>%
  mutate(perc_growth_Jan=((Jan-Oct)/Oct)*100)%>%
  mutate(perc_growth_Feb=((Feb-Oct)/Oct)*100)%>%
  mutate(perc_growth_April=((April-Oct)/Oct)*100)%>%
  mutate(perc_growth_May=((May-Oct)/Oct)*100)
```

### Plot growth for adult animals sampled in November

```{r}
plot3<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Nov)%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Nov, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Nov Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot3

ggsave(plot=plot3, "lifestage_carryover/output/size/adults_Nov_percent_growth.png", width=4, height=4)
```

Test for effect of treatment. 
```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Nov)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Nov~treatment, data=.)

summary(model)
```
Effect of treatment just barely not significant (P=0.088). 

            Df Sum Sq Mean Sq F value Pr(>F)  
treatment    1  8.358   8.358   4.132 0.0883 .
Residuals    6 12.137   2.023    

### Plot growth for animals measured in January 

Adults
```{r}
plot4<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="adult")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  ylim(-7, 3)+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Jan Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot4

ggsave(plot=plot4, "lifestage_carryover/output/size/adults_Jan_percent_growth.png", width=4, height=4)
```

Juveniles
```{r}
plot5<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="juvenile")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Juvenile: Oct-Jan Growth")+
  theme_classic()+
  ylim(0,20)+
  theme(
    text=element_text(size=12, color="black")
  );plot5

ggsave(plot=plot5, "lifestage_carryover/output/size/juvenile_Jan_percent_growth.png", width=4, height=4)
```

Seed

```{r}
plot6<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="seed")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Seed: Oct-Jan Growth")+
  theme_classic()+
  ylim(0,250)+
  theme(
    text=element_text(size=12, color="black")
  );plot6

ggsave(plot=plot6, "lifestage_carryover/output/size/seed_Jan_percent_growth.png", width=4, height=4)
```

Spat

```{r}
plot7<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-Jan Growth")+
  theme_classic()+
  ylim(-10,1)+
  theme(
    text=element_text(size=12, color="black")
  );plot7

ggsave(plot=plot7, "lifestage_carryover/output/size/spat_Jan_percent_growth.png", width=4, height=4)
```


### Plot growth for animals measured in February 

Adults
```{r}
plot8<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="adult")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  ylim(-7, 3)+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Feb Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot8

ggsave(plot=plot8, "lifestage_carryover/output/size/adults_Feb_percent_growth.png", width=4, height=4)
```

Juveniles
```{r}
plot9<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="juvenile")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Juvenile: Oct-Feb Growth")+
  theme_classic()+
  #ylim(0,20)+
  theme(
    text=element_text(size=12, color="black")
  );plot9

ggsave(plot=plot9, "lifestage_carryover/output/size/juvenile_Feb_percent_growth.png", width=4, height=4)
```

Seed
```{r}
plot10<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="seed")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Seed: Oct-Feb Growth")+
  theme_classic()+
  ylim(0,250)+
  theme(
    text=element_text(size=12, color="black")
  );plot10

ggsave(plot=plot10, "lifestage_carryover/output/size/seed_Feb_percent_growth.png", width=4, height=4)
```

Spat
```{r}
plot11<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-Feb Growth")+
  theme_classic()+
  ylim(-10,15)+
  theme(
    text=element_text(size=12, color="black")
  );plot11

ggsave(plot=plot11, "lifestage_carryover/output/size/spat_Feb_percent_growth.png", width=4, height=4)
```


### Plot growth for animals measured in April (spat only)

Spat
```{r}
plot11b<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_April)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_April, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-April Growth")+
  theme_classic()+
  ylim(-10,15)+
  theme(
    text=element_text(size=12, color="black")
  );plot11b

ggsave(plot=plot11b, "lifestage_carryover/output/size/spat_April_percent_growth.png", width=4, height=4)
```



### Plot growth for animals measured in May (spat only)

Spat
```{r}
plot11c<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_May)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_May, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-May Growth")+
  theme_classic()+
  ylim(-10,15)+
  theme(
    text=element_text(size=12, color="black")
  );plot11c

ggsave(plot=plot11c, "lifestage_carryover/output/size/spat_May_percent_growth.png", width=4, height=4)
```


### Test for effect of treatment by lifestage 

January all stages 

```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Jan~treatment*lifestage, data=.)

summary(model)
```

Effect of lifestage, no treatment effects. 

                    Df Sum Sq Mean Sq F value Pr(>F)    
treatment            1     22      22   0.223  0.644    
lifestage            3  94337   31446 319.452  4e-13 ***
treatment:lifestage  3     47      16   0.160  0.921    
Residuals           14   1378      98    


February all stages 

```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Feb~treatment*lifestage, data=.)

summary(model)
```

Treatment x lifestage effects 

                    Df Sum Sq Mean Sq F value   Pr(>F)    
treatment            1    129     129   1.185    0.295    
lifestage            3  81243   27081 249.636 2.18e-12 ***
treatment:lifestage  3   1683     561   5.172    0.013 *  
Residuals           14   1519     108   


## Plot percent growth in only ambient treatment 

Plot growth rates between lifestages. 

Calculate growth as % growth per day. 

```{r}
replicates<-replicates%>%
  mutate(perc_growth_day=perc_growth_Feb/90)
```


```{r}
plot12<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_day)%>%
  #filter(treatment=="control")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(x=lifestage, y=perc_growth_day))+
  geom_point()+
  geom_boxplot(fill=NA)+
  ylab("% Winter Growth per Day")+
  xlab("Life Stage \nLength")+
  scale_x_discrete(limits=c("spat", "seed", "juvenile", "adult"), labels=c("Spat \n<15mm", "Seed \n15-25mm", "Juvenile \n25-40mm", "Adult \n>40mm"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black"), 
    axis.text=element_text(size=12, color="black"), 
    axis.title=element_text(size=14, color="black", face="bold")
  );plot12

ggsave(plot=plot12, "lifestage_carryover/output/size/lifestage_growth.png", width=5, height=5)
```

Calculate size of each lifestage 
```{r}
avg_size<-replicates%>%
  select(lifestage, Feb)%>%
  group_by(lifestage)%>%
  summarise(mean=mean(Feb, na.rm=TRUE));avg_size
```

```{r}
avg_size<-replicates%>%
  select(lifestage, tag, Feb)%>%
  filter(lifestage=="adult")%>%
  group_by(tag)%>%
  summarise(mean=mean(Feb, na.rm=TRUE));avg_size
```

# Individual growth tracking phase (Adults, seed, and juveniles)   

Load data for individual growth measurements

```{r}
individuals<-read_csv("lifestage_carryover/data/size/gigas-lengths-individuals.csv")

individuals<-individuals%>%
  select(!Notes)

individuals$unique_id<-paste0(individuals$bag, "-", individuals$oyster)
```

Calculate individual growth by calculating the percent change from April to May for each individual. We will not use March due too a caliper error.  

```{r}
individuals_wide<-individuals%>%
  mutate(perc_growth_May=((April-May)/April)*100)

individuals_long<-individuals%>%
  pivot_longer(names_to="month", values_to="length.mm", cols=c(March:May))
```

# Plot and run anova models  

Plot mean lengths for adults and seed. Remove March.  

```{r}
plot_ind<-individuals_long%>%
  filter(!month=="March")%>%
  
  ggplot(aes(color=`init-treat`, x=month, y=length.mm))+
  geom_point(size=1)+
  geom_line(aes(group=unique_id))+
  facet_grid(~`life-stage`)+
  ylab("Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot_ind

ggsave(plot=plot_ind, "lifestage_carryover/output/size/individual_length_time.png", width=7, height=4)
```

Plot adults by treatment and family with percent growth. 

```{r}
plot_ind1<-individuals_wide%>%
  filter(`life-stage`=="adult")%>%
  
  ggplot(aes(color=`init-treat`, x=tag, y=perc_growth_May))+
  geom_boxplot(fill=NA, outlier.shape=NA)+
  geom_point(position=position_dodge(0.7), size=1)+
  ylab("Growth (%)")+
  xlab("Family")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot_ind1

ggsave(plot=plot_ind1, "lifestage_carryover/output/size/individual_growth_adults.png", width=7, height=4)
```

Plot seed by treatment with percent growth. 

```{r}
plot_ind2<-individuals_wide%>%
  filter(`life-stage`=="seed")%>%
  
  ggplot(aes(color=`init-treat`, x=`init-treat`, y=perc_growth_May))+
  geom_boxplot(fill=NA, outlier.shape=NA)+
  geom_point(position=position_dodge(0.7), size=1)+
  ylab("Growth (%)")+
  xlab("Treatment")+
  ylim(-40,10)+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot_ind2

ggsave(plot=plot_ind2, "lifestage_carryover/output/size/individual_growth_seed.png", width=7, height=4)
```

# Run ANOVA on individual growth 

Adults 
```{r}
model<-individuals_wide%>%
  filter(`life-stage`=="adult")%>%
  
  aov(perc_growth_May ~ `init-treat`, data=.)

summary(model)
```
No difference in adults.  

Seed 
```{r}
model<-individuals_wide%>%
  filter(`life-stage`=="seed")%>%
  
  aov(perc_growth_May ~ `init-treat`, data=.)

summary(model)
```
No difference in seed.    

