---
title: "C. gigas growth - lifestage carryover project" 
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load libraries 

```{r}
library(tidyverse)
library(ggplot2)
```

# Read in data 

```{r}
growth<-read_csv("lifestage_carryover/data/size/gigas-lengths.csv")

growth<-growth%>%
  rename(tag=`tag...1`)%>%
  select(tag, `life-stage`, `init-treat`, label, month, length)%>%
  rename(lifestage=`life-stage`, treatment=`init-treat`)

growth$month<-factor(growth$month, levels=c("Oct", "Nov", "Jan", "Feb"))
```

# Look at data for whole population together  

## Plot growth over months for each life stage and treatment 

Plot as box plot. Filter out adults that were sampled in November, since these were larger and are not present in later dates. 
```{r}
plot<-growth%>%
  filter(!tag=="Cat")%>%
  filter(!tag=="Dog")%>%
  filter(!tag=="Yellow")%>%
  filter(!tag=="Toad")%>%
  
  ggplot(aes(color=treatment, x=month, y=length))+
  geom_boxplot(fill=NA, outlier.shape=NA)+
  geom_point(position=position_dodge(0.7), size=1)+
  facet_grid(~lifestage)+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot

ggsave(plot=plot, "lifestage_carryover/output/size/mean_lengths_box.png", width=7, height=4)
```

Plot as mean +- standard error for all.  
```{r}
plot2<-growth%>%
  filter(!tag=="Cat")%>%
  filter(!tag=="Dog")%>%
  filter(!tag=="Yellow")%>%
  filter(!tag=="Toad")%>%
  
  group_by(treatment, month, lifestage)%>%
  summarise(mean=mean(length, na.rm=TRUE), se=sd(length, na.rm=TRUE)/sqrt(length(length)))%>%
  
  ggplot(aes(color=treatment, x=month, y=mean))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=1)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, position=position_dodge(0.3), linewidth=0.75)+
  geom_line(aes(group=treatment), position=position_dodge(0.3), linewidth=1)+
  facet_grid(~lifestage)+
  ylab("Mean Length (mm)")+
  xlab("Month")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot2

ggsave(plot=plot2, "lifestage_carryover/output/size/mean_lengths.png", width=7, height=4)
```

## Run model

```{r}
model<-aov(length~lifestage*month*treatment, data=growth)
summary(model)
```

                            Df Sum Sq Mean Sq  F value  Pr(>F)    
lifestage                    3 521245  173748 2201.287 < 2e-16 ***
month                        3  13168    4389   55.610 < 2e-16 ***
treatment                    1      2       2    0.025   0.874    
lifestage:month              6  22776    3796   48.093 < 2e-16 ***
lifestage:treatment          3   1681     560    7.098 9.8e-05 ***
month:treatment              3     76      25    0.321   0.810    
lifestage:month:treatment    6    466      78    0.983   0.435    
Residuals                 1490 117606      79  

Conduct posthoc test
```{r}
library(emmeans)
emm<-emmeans(model, ~ treatment | month | lifestage)
pairs(emm)
```

Growth is different in February for spat. No other differences.  

# Look at growth as percent growth summarised at the level of replicate containers or tags 

Calculate mean size at each month for each replicate and calculate % growth (change/original).  

```{r}
replicates<-growth%>%
  group_by(treatment, month, lifestage, tag)%>%
  summarise(mean=mean(length, na.rm=TRUE))%>%
  pivot_wider(names_from=month, values_from=mean)%>%
  mutate(perc_growth_Nov=((Nov-Oct)/Oct)*100)%>%
  mutate(perc_growth_Jan=((Jan-Oct)/Oct)*100)%>%
  mutate(perc_growth_Feb=((Feb-Oct)/Oct)*100)
```

## Plot growth for adult animals sampled in November

```{r}
plot3<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Nov)%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Nov, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Nov Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot3

ggsave(plot=plot3, "lifestage_carryover/output/size/adults_Nov_percent_growth.png", width=4, height=4)
```

Test for effect of treatment. 
```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Nov)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Nov~treatment, data=.)

summary(model)
```
Effect of treatment just barely not significant (P=0.088). 

            Df Sum Sq Mean Sq F value Pr(>F)  
treatment    1  8.358   8.358   4.132 0.0883 .
Residuals    6 12.137   2.023    

## Plot growth for animals measured in January 

Adults
```{r}
plot4<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="adult")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  ylim(-7, 3)+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Jan Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot4

ggsave(plot=plot4, "lifestage_carryover/output/size/adults_Jan_percent_growth.png", width=4, height=4)
```

Juveniles
```{r}
plot5<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="juvenile")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Juvenile: Oct-Jan Growth")+
  theme_classic()+
  ylim(0,20)+
  theme(
    text=element_text(size=12, color="black")
  );plot5

ggsave(plot=plot5, "lifestage_carryover/output/size/juvenile_Jan_percent_growth.png", width=4, height=4)
```

Seed

```{r}
plot6<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="seed")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Seed: Oct-Jan Growth")+
  theme_classic()+
  ylim(0,250)+
  theme(
    text=element_text(size=12, color="black")
  );plot6

ggsave(plot=plot6, "lifestage_carryover/output/size/seed_Jan_percent_growth.png", width=4, height=4)
```

Spat

```{r}
plot7<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Jan, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-Jan Growth")+
  theme_classic()+
  ylim(0,150)+
  theme(
    text=element_text(size=12, color="black")
  );plot7

ggsave(plot=plot7, "lifestage_carryover/output/size/spat_Jan_percent_growth.png", width=4, height=4)
```



## Plot growth for animals measured in February 

Adults
```{r}
plot8<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="adult")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  ylim(-7, 3)+
  scale_shape_manual(name="Family", values=c(15, 16, 17, 18))+
  ggtitle("Adults: Oct-Feb Growth")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot8

ggsave(plot=plot8, "lifestage_carryover/output/size/adults_Feb_percent_growth.png", width=4, height=4)
```

Juveniles
```{r}
plot9<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="juvenile")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Juvenile: Oct-Feb Growth")+
  theme_classic()+
  ylim(0,20)+
  theme(
    text=element_text(size=12, color="black")
  );plot9

ggsave(plot=plot9, "lifestage_carryover/output/size/juvenile_Feb_percent_growth.png", width=4, height=4)
```

Seed
```{r}
plot10<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="seed")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Seed: Oct-Feb Growth")+
  theme_classic()+
  ylim(0,250)+
  theme(
    text=element_text(size=12, color="black")
  );plot10

ggsave(plot=plot10, "lifestage_carryover/output/size/seed_Feb_percent_growth.png", width=4, height=4)
```

Spat
```{r}
plot11<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  filter(lifestage=="spat")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(color=treatment, x=treatment, y=perc_growth_Feb, shape=tag))+
  #geom_line(aes(group=tag), color="black")+
  geom_point(aes(group=interaction(treatment, lifestage)), position=position_dodge(0.3), size=5)+
  ylab("% growth")+
  xlab("Treatment")+
  scale_color_manual(values=c("darkgray", "darkred"), name="Treatment", labels=c("Control", "Treatment"))+
  scale_shape_manual(name="Replicate", values=c(15, 16, 17, 18))+
  ggtitle("Spat: Oct-Feb Growth")+
  theme_classic()+
  ylim(0,250)+
  theme(
    text=element_text(size=12, color="black")
  );plot11

ggsave(plot=plot11, "lifestage_carryover/output/size/spat_Feb_percent_growth.png", width=4, height=4)
```

## Test for effect of treatment by lifestage 

January all stages 

```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Jan)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Jan~treatment*lifestage, data=.)

summary(model)
```

Effect of lifestage, no treatment effects. 

                    Df Sum Sq Mean Sq F value   Pr(>F)    
treatment            1      1       1   0.015    0.904    
lifestage            3  96056   32019 325.275 3.54e-13 ***
treatment:lifestage  3    363     121   1.231    0.335    
Residuals           14   1378      98   


February all stages 

```{r}
model<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_Feb)%>%
  drop_na()%>%
  droplevels()%>%
  
  aov(perc_growth_Feb~treatment*lifestage, data=.)

summary(model)
```

Treatment x lifestage effects 

                    Df Sum Sq Mean Sq F value   Pr(>F)    
treatment            1    393     393   3.622   0.0778 .  
lifestage            3 109285   36428 335.801 2.84e-13 ***
treatment:lifestage  3  10028    3343  30.814 2.01e-06 ***
Residuals           14   1519     108   

# Plot percent growth in only ambient treatment 

Plot growth rates between lifestages. 

Calculate growth as % growth per day. 

```{r}
replicates<-replicates%>%
  mutate(perc_growth_day=perc_growth_Feb/90)
```


```{r}
plot12<-replicates%>%
  select(treatment, lifestage, tag, perc_growth_day)%>%
  #filter(treatment=="control")%>%
  drop_na()%>%
  droplevels()%>%
  
  ggplot(aes(x=lifestage, y=perc_growth_day))+
  geom_point()+
  geom_boxplot(fill=NA)+
  ylab("% Winter Growth per Day")+
  xlab("Life Stage \nLength")+
  scale_x_discrete(limits=c("spat", "seed", "juvenile", "adult"), labels=c("Spat \n<15mm", "Spat \n15-25mm", "Seed \n25-40mm", "Seed \n>40mm"))+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black"), 
    axis.text=element_text(size=12, color="black"), 
    axis.title=element_text(size=14, color="black", face="bold")
  );plot12

ggsave(plot=plot12, "lifestage_carryover/output/size/lifestage_growth.png", width=5, height=5)
```

Calculate size of each lifestage 
```{r}
avg_size<-replicates%>%
  select(lifestage, Feb)%>%
  group_by(lifestage)%>%
  summarise(mean=mean(Feb, na.rm=TRUE));avg_size
```

```{r}
avg_size<-replicates%>%
  select(lifestage, tag, Feb)%>%
  filter(lifestage=="adult")%>%
  group_by(tag)%>%
  summarise(mean=mean(Feb, na.rm=TRUE));avg_size
```
