---
title: "Resazurin trials - lifestage carryover project" 
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load libraries 

```{r}
library(tidyverse)
library(ggplot2)
```

# Read in data 

```{r}
data<-read_csv("lifestage_carryover/data/resazurin_trial_02/20240312-cgig-resazurin-trial-02.csv")

data<-data%>%
  select(well, sample, type, `weight-mg`, treatment, `t0`, `t0.5`, `t1.0`)%>%
  rename(weight.mg=`weight-mg`, time0=`t0`, time0.5=`t0.5`, time1=`t1.0`)

```

# Calculate corrected values

Calculate change in 0.5 and 1 time values from the initial t0 value. 
```{r}

data<-data%>%
  mutate(change_time0.5=time0.5-time0)%>%
  mutate(change_time1=time1-time0)
```

Calculate average of change in negative controls for pre-stress and stress plates for normalization. 

```{r}
neg_stress<-data%>%
  filter(type=="neg-control" & treatment=="heat_stress")%>%
  group_by(treatment)%>%
  summarise(neg_stress_mean0.5=mean(change_time0.5), neg_stress_mean1=mean(change_time1))

neg_prestress<-data%>%
  filter(type=="neg-control" & treatment=="pre-stress")%>%
  group_by(treatment)%>%
  summarise(neg_control_mean0.5=mean(change_time0.5), neg_control_mean1=mean(change_time1))
```

Add values into data frame for controls. 

```{r}
data<-data%>%
  mutate(neg_stress_mean0.5=neg_stress$neg_stress_mean0.5)%>%
  mutate(neg_stress_mean1=neg_stress$neg_stress_mean1)%>%
  mutate(neg_control_mean0.5=neg_prestress$neg_control_mean0.5)%>%
  mutate(neg_control_mean1=neg_prestress$neg_control_mean1)

head(data)
```

Subtract changes in negative controls from changes in oyster values to generate a corrected value. 

```{r}
data<-data%>%
  mutate(corr_time0.5=if_else(treatment=="pre-stress", change_time0.5-neg_control_mean0.5, 
                              if_else(treatment=="heat_stress", change_time0.5-neg_stress_mean0.5, NA)))%>%
  
  mutate(corr_time1=if_else(treatment=="pre-stress", change_time1-neg_control_mean1, 
                              if_else(treatment=="heat_stress", change_time1-neg_stress_mean1, NA)))
  
```

Select data we want to keep. 

```{r}
data<-data%>%
  select(well, sample, type, weight.mg, treatment, corr_time1)%>%
  filter(!type=="neg-control")%>%
  filter(!type=="empty")
```

There are some negative values at time 0.5, so lets just use time 1 for now (end of run).  

Pivot wider. 

```{r}
data<-data%>%
  pivot_wider(names_from=treatment, values_from=corr_time1)
```

Calculate percent change in values after stress compared to control. 

```{r}
data<-data%>%
  mutate(perc_change=100*(heat_stress-`pre-stress`)/`pre-stress`)
```

# Plot data scaled by weight separate for treatments  

Plot percent change. 
```{r}
plot1<-data%>%
  filter(perc_change<3000)%>% #remove outlier
  filter(weight.mg<2000)%>% #remove large oyster
  
  ggplot(aes(x=weight.mg, y=perc_change))+
  #geom_smooth(color="darkred", se=FALSE, method="lm")+
  geom_vline(xintercept=100)+
  geom_hline(yintercept=0)+
  geom_point(size=3)+
  ylab("Percent Change [Stress-Control]")+
  xlab("Oyster")+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot1
```

Plot only control readings. 
```{r}
plot2<-data%>%
  #filter(perc_change<3000)%>% #remove outlier
  filter(weight.mg<2000)%>% #remove large oysters
  filter(heat_stress<7500)%>%
  
  ggplot(aes(x=weight.mg, y=`pre-stress`))+
  geom_smooth(color="darkgray", se=FALSE, method="lm")+
  geom_point(size=3)+
  ylab("Fluorescence [Control]")+
  xlab("Weight (mg)")+
  ylim(0,15000)+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot2
```

Plot only stress readings. 
```{r}
plot3<-data%>%
  #filter(perc_change<3000)%>% #remove outlier
  filter(weight.mg<2000)%>% #remove large oysters
  filter(heat_stress<7500)%>%
  
  ggplot(aes(x=weight.mg, y=heat_stress))+
  geom_smooth(color="darkred", se=FALSE, method="lm")+
  geom_point(size=3)+
  ylab("Fluorescence [Stress]")+
  xlab("Weight (mg)")+
  ylim(0,15000)+
  theme_classic()+
  theme(
    text=element_text(size=12, color="black")
  );plot3
```

Generate plot for grant. 

```{r}
library(cowplot)

plots<-plot_grid(plot2, plot3, ncol=2, nrow=1, labels = c('B', 'C'), align="vh")

ggsave(filename="lifestage_carryover/output/resazurin/scale_weight.png", plot=plots, dpi=300, width=6, height=4, units="in")
```

# Plot data scaled by weight together for treatments  

Re arrange data to long format. 

```{r}
long<-data%>%
  select(!perc_change)%>%
  pivot_longer(cols=c("pre-stress", "heat_stress"), values_to="value", names_to="treatment")
```

Plot. 

```{r}
plot4<-long%>%
  filter(weight.mg<400)%>% #remove large oysters
  filter(!c(treatment=="heat_stress" & value>7500))%>% #remove outliers as done above
  
  ggplot(aes(x=weight.mg, y=value, color=treatment))+
  geom_smooth(aes(group=treatment),se=FALSE, method="lm")+
  geom_point(size=3)+
  scale_color_manual(values=c("darkred", "darkgray"), labels=c("heat stress", "control"), name="Treatment")+
  ylab("Change in Fluorescence (1 hr)")+
  xlab("Weight (mg)")+
  ylim(0,15000)+
  theme_classic()+
  ggtitle("B")+
  theme(
    text=element_text(size=12, color="black")
  );plot4

ggsave(filename="lifestage_carryover/output/resazurin/treatment_scale_weights.png", plot=plot4, dpi=300, width=5, height=4, units="in")
```




